---
title: "Homework 5"
author: "Hyunjoon Rhee"
date: "3/28/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Problem 1
### a)
```{r}
library(faraway)
data(aatemp)
prob1 = lm(temp ~ year, data = aatemp)
plot(aatemp$year, aatemp$temp)
abline(prob1)
cor(aatemp$year, aatemp$temp)
```
The graph shows that there is a weak linear relationship between year and the temperature. The correlation of 0.292 shows the linear trend.

### b)
```{r}
library(lmtest)
dwtest(prob1)
```

Because the dwtest shows that it has a p value that is smaller than 0.05, reject the null hypothesis meaning that it can be said that there is correlation in the error. 
```{r}
library(nlme)
prob1_b = gls(temp ~ year, correlation = corAR1(form= ~year), data=aatemp)
summary(prob1_b)
```
Although the Phi Coefficient is not significant with a value of 0.2303, the residual standard error of the model that is fitted with autocorrelated error is 1.475, which means that there can be a possibility of a linear fit.

### c)
```{r}
prob1_c = lm(temp ~ I(year) + I(year^2) + I(year^3) + I(year^4) + I(year^5) + I(year^6) + I(year^7) +I(year^8) + I(year^9) + I(year^10), data=aatemp)
prob1_back = step(prob1_c, direction = 'backward', trace=10)
summary(prob1_back)
prob1_c2 = lm(temp ~ I(year) + I(year^2) + I(year^3) + I(year^4) + I(year^8), data=aatemp)
```

```{r}
library(ggplot2)
ggplot(data = aatemp, aes(x = year, y = temp)) + geom_point() + geom_smooth(method = 'lm', formula = y ~ x + I(x^2) + I(x^3) + I(x^4) + I(x^8)) + labs(x = 'Year', y = 'Temp')
predict(prob1_back, newdata = data.frame(year=2001))
```
The model fitted shows a pattern of a linear model. The predicted value at year 2001 is 49.29922

### d)
```{r}
library(splines)
prob1_d = lm(temp~bs(year, df = 6, intercept = TRUE), data = aatemp)
ggplot(data = aatemp, aes(x = year, y = temp)) + geom_point() + geom_smooth(method = 'lm', formula = y ~ bs(x, df = 6), aes(color = 'Spline')) + geom_smooth(method = 'lm', linetype = 'dashed', formula = y ~ x + I(x^2) + I(x^3) + I(x^4) + I(x^8), aes(color = ('Polynomial'))) + labs(x = 'Year', y = 'Temp')
```
The plot above shows that the two models are almost equal, which means that there is no much better fitted model. But just by looking at the data points, it still seems to have a linear relationship.

## Problem 2
### a)
```{r}
data(infmort)
head(infmort)
```

Income and mortality are numerical variables and region and oil are categorical variables.

### b)
```{r}
prob2_b = lm(mortality ~ ., data = infmort)
summary(prob2_b)
```

```{r}
prob2_b2 = lm(mortality ~ region + income + oil + region*income + region*oil + income*oil, data = infmort)
summary(prob2_b2)
```

```{r}
plot(prob2_b2)
```
The model seems to show normal distribution and constant variance. Model could remove Afganistan and Saudi Arabia from the data to show better data, but it cannot be assumed that it has a linear relationship.

### c)
```{r}
library(MASS)
prob2_b3 = lm(mortality ~ region + log(income) + oil + region*log(income) + region*oil + income*oil, data = infmort)
summary(prob2_b3)
plot(prob2_b3)
```
log transformation could be performed. After using log transformation the data seems to have better constant variance. But there is not much difference.

### d)
In Asia, America, Asia*oil export, the mortality is decreased, but rest of them increases the mortality.