---
title: "Homework 6"
author: "Hyunjoon Rhee"
date: "5/3/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Problem 1
### a)
```{r}
library(faraway)
attach(butterfat)
```

```{r}
head(butterfat)
```

```{r}
par(mfrow=c(1,2))
plot(Breed, Butterfat, data = butterfat)
plot(Age, Butterfat, data = butterfat)
```

### b)
```{r}
model=aov(Butterfat~Breed+Age+Breed*Age, data = butterfat)
summary(model)
```
Because the p value for breed and age interaction term is greater than $\alpha = 0.05$ which means that it fails to reject that there is an interaction term has zero effect on the response.

### c)
```{r}
maineffect = lm(Butterfat ~ Breed + Age, data = butterfat)
summary(maineffect)
```
From problem 1 b) it was concluded that the interaction term has no statistical meaning. Therefore, the additive model was introduced. Looking at the p value of each term, only the 'Age' term has a p value that is greater than $\alpha = 0.05$. This means that only age rejects the null hypothesis and conclude that age does not have statistical significance. 

```{r}
anova(lm(Butterfat ~ Breed, data=butterfat))
anova(lm(Butterfat ~ Age, data=butterfat))
```
Looking at the anova table of the two of the p values, it is evident that there is a statistical difference between the breed and the butterfat, whereas age does not have statistical difference between age and butterfat. This is the same result as the above analysis.

### d)
Constant variance check
```{r}
plot(maineffect$fitted.values, maineffect$residuals)
library(lmtest)
bptest(maineffect)
```
It is hard to see from the graph if there is a constant variance, so Breusch-Pagan test was performed. The bptest shows a p value that is less than $\alpha = 0.05$, which means it can reject the null hypothesis and conclude that it is not homoskedastic.

Normality check
```{r}
qqnorm(maineffect$residuals)
qqline(maineffect$residuals)
shapiro.test(maineffect$residuals)
```
The qqplot seems to show a straight line, but for further analysis, shapiro-wilk test was performed. Because the p value is smaller than $\alpha=0.05$, it is evident that it can reject the null hypothesis and conclude that the residual is not normally distributed. 

```{r}
dwtest(maineffect)
```
The Durbin-Watson test shows it has a p value that is greater than $\alpha=0.05$, which means that it fails to reject the null hypothesis and conclude that the errors are statistically not correlated.

Constant variance ,normality, correlation assumptions were not met, meaning that them model itself is questionable.

### e)
```{r}
library(tidyverse)
bestbreed = butterfat %>% filter(Breed == 'Jersey' | Breed == 'Guernsey') %>% select(Butterfat, Breed)
t.test(Butterfat ~ Breed, data = bestbreed)
```
Looking at the results in Problem 1 a), it is known that Jersey and Guernsey are the two breeds that has most butterfat content. Because the p value of the t test is greater than $\alpha=0.05$, it fails to reject the null hypothesis and conclude that it is hard to say there is a statistical difference between the best and the second.


## Problem 2
### a)
```{r}
anova(lm(Speed ~ Run + Expt, data = morley))
```

```{r}
anova(lm(Speed ~ Run, data = morley))
```
Looking at the p value of the anova table, it is much greater than $\alpha=0.05$, which means that there is not a significant difference among the run groups.

### b)
By using blocking factor, it can categorize the difference between the run groups and compare the response. Without a blocking factor, it would have not earned any meaningful data related to the mean of the speed.


## Problem 3
### a)
```{r}
prob3 = lm(yield~ ., alfalfa)
anova(prob3)
```
The anova table shows that when the $\alpha=0.05$, shade and inoculum show statistical significance whereas irrigation does not.

### b)
```{r}
TukeyHSD(aov(yield~., alfalfa), "inoculum")
```
The confidence interval shows that everything but E-A, E-B, E-C, E-D do not contain 0 in the confidence interval, which means that there is no significant difference between every pair but the ones that contains E in it. This shows that A-D show significant difference with E, but the rest do not. 
