---
title: "HW3"
author: "Hyunjoon Rhee"
date: "3/14/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

Problem 1
```{r}
library(faraway)
data(salmonella)
head(salmonella)
```
a)
```{r}
a = lm(colonies~log(dose+1),data=salmonella)
summary(a)
```

b)
```{r}
plot(a,which=1)
```
The first plot shows the residual vs fitted.

c)
```{r}
c = lm(colonies~factor(log(dose+1)),data=salmonella)
anova(a,c)
```
Because the p value is greater than $\alpha = 0.05$ it fails to reject the null, which means that it has no lack of fit.

Problem 2
```{r}
library(faraway)
library(lmtest)
data("gammaray")
prob2 = lm(flux ~ time, data = gammaray)
bptest(prob2)
dwtest(prob2)
plot(prob2,which=1)
```

a)
```{r}
prob21 = lm(log(flux) ~ log(time), data = gammaray)
prob22 = lm(log(flux) ~ time, data = gammaray)
prob23 = lm(flux ~ log(time), data = gammaray)
prob24 = lm(flux ~ time, data = gammaray)
par(mfrow=c(2,2))
plot(prob21,which=1)
plot(prob22,which=1)
plot(prob23,which=1)
plot(prob24,which=1)
```
```{r}
prob2a = lm(log(flux) ~ log(time), data = gammaray)
bptest(prob2a)
dwtest(prob2a)
plot(prob2a,which=1)
```
Transformation of log for both seems to show better.

b)
```{r}
prob2b = lm(log(flux)~log(time), data=gammaray,weights = 1/abs(error))
```

c)
```{r}
summary(prob2b)
bptest(prob2b)
shapiro.test(residuals(prob2b))
```
Because the Breusch-Pagan test shows that it has p value less than 0.05, it has constant variance. Also because Shapiro test value is less than 0.05, it can be assumed that it is normal.

d)
I believe this log transformation on both end is the best model because it both meets constant variance and normality assumption.

Problem 3
a)
```{r}
data(divusa)
head(divusa)
prob3 = lm(divorce ~ unemployed+femlab+marriage+birth+military, data = divusa)
plot(prob3,which=1)

```
It is hard to say that it is autocorrelated just by looking at the graph, but I believe it is autocorrelated because it looks like it is connected from one after another.

b)
```{r}
dwtest(prob3)
```
Null hypothesis is that errors are not autocorrelated. However, since the pvalue is smaller than 0.05, we reject the null hypothesis, meaning that errors are autocorrelated.

c)
```{r}
library(nlme)
prob3c = gls(divorce ~ unemployed+femlab+marriage+birth+military, correlation= corAR1(form=~year),method='ML',data=divusa)
summary(prob3c)
```
d)
```{r}
intervals(prob3c)
```
Because the confidence interval does not contain 0, we reject the null hypothesis. Therefore correlation coefficient is significant.

e)
GLS model does not change the variables that are shown significant in the lm model.

f)
Autocorrelation gives a potential reason why there might be a error in correlation, due to its time sequential properties.
